\documentclass[12pt]{beamer} %Makes presentation
\input{preamble}

\title{Preview of\\``Public opinion matters?''}

\date[]{November 13, 2013}

\begin{document}

\frame{\titlepage}

\frame{
\heading{Overview for next week}
}


\frame<4>[label=readings]{
\heading{Readings for Next Week}
\begin{itemize}\itemsep1em
\item<1-> Erikson -- ``Constituency Opinion and Congressional Behavior''
\item<2-> Stimson et al. -- ``Dynamic Representation''
\item<3-> Jacobs and Page ``Who Influences U.S. Foreign Policy?''
\item<4-> Riker -- ``Liberalism Against Populism'' (from last week)
\end{itemize}
}

\againframe<1>{readings}

\frame{
\heading{Erikson}
\begin{itemize}\itemsep1em
\item Skip this. I'll talk about it without you having to read the details.
\end{itemize}
}

\againframe<2>{readings}

\frame{
\heading{Stimson et al.}
\begin{itemize}\itemsep1em
\item How can we measure whether opinion is being represented in policy?
\item Does it make sense to use ``macropartisanship'' to measure representation?
\item Don't worry about American politics details, focus on theory and method.
\end{itemize}
}

\againframe<3>{readings}

\frame{
\heading{Jacobs and Page}
\begin{itemize}\itemsep1em
\item What happens if we don't treat ``public opinion'' as a singular entity?
\item Whose opinions get represented?
\end{itemize}
}

\againframe<4>{readings}

\frame{
\heading{Riker}
\begin{itemize}\itemsep1em
\item Is representative democracy necessary? Or could we have direct democracy?
\item Depending on the answer, what does that mean for how we should think about representation?
\item This is formal theoretic, but hopefully is accessible
\end{itemize}
}


\appendix
\frame{}

\end{document}
