\documentclass[12pt]{beamer} %Makes presentation
%\documentclass[handout]{beamer} %Makes Handouts
\usetheme{Singapore} %Gray with fade at top
\useoutertheme[subsection=false]{miniframes} %Supppress subsection in header
\useinnertheme{rectangles} %Itemize/Enumerate boxes
\usecolortheme{seagull} %Color theme
\usecolortheme{rose} %Inner color theme

\definecolor{light-gray}{gray}{0.75}
\definecolor{dark-gray}{gray}{0.55}
\setbeamercolor{item}{fg=light-gray}
\setbeamercolor{enumerate item}{fg=dark-gray}

\setbeamertemplate{navigation symbols}{}
\setbeamertemplate{mini frames}[default]
\setbeamercovered{dynamics}
\setbeamerfont*{title}{size=\Large,series=\bfseries}

%\setbeameroption{notes on second screen} %Dual-Screen Notes
%\setbeameroption{show only notes} %Notes Output

\newcommand{\heading}[1]{\noindent \textbf{#1}\\ \vspace{1em}}

\usepackage{bbding,color,multirow,times,ccaption,tabularx,graphicx,verbatim,booktabs,fixltx2e}
\usepackage{colortbl} %Table overlays
\usepackage[english]{babel}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}

\title{Preview of\\``How do people form opinions?''}

%\author[]{Thomas J. Leeper}
\institute[]{
  \inst{}%
  Department of Political Science and Government\\Aarhus University
}
\date[]{September 11, 2013}

\begin{document}

\frame{\titlepage}

\frame{
\heading{Overview for next week}
}


\frame<4>[label=readings]{
\heading{Readings for Next Week}
\begin{itemize}\itemsep1em
\item<1-> Eagly and Chaiken -- ``Attitude Structure and Function''
\item<2-> Druckman and Lupia -- ``Preference Formation''
\item<3-> Zaller -- {\em Nature and Origins of Mass Opinion}
\item<4-> Lau and Redlawsk -- ``Advantages and Disadvantages of Cognitive Heuristics''
\end{itemize}
}

\againframe<1>{readings}

\frame{
\heading{Eagly and Chaiken}
\begin{itemize}\itemsep1em
\item Note: Skip section on behavior (pp.295--303)
\item Psychological foundations
\item What is an attitude? How is it structured?
\item What function(s) do attitudes serve?
\item What makes attitudes strong?
\end{itemize}
}

\againframe<2>{readings}

\frame{
\heading{Druckman and Lupia}
\begin{itemize}\itemsep1em
\item What is a political preference?
\item Are preferences and attitudes equivalent?
\item How do people form attitudes/preferences?
\item Do people in general have preferences?
\end{itemize}
}

\againframe<3>{readings}

\frame{
\heading{Zaller}
\begin{itemize}\itemsep1em
\item How does Zaller's model compare to the other models described by Druckman and Lupia?
\item Is Zaller talking about attitudes or preferences?
\item Is Zaller talking about attitude formation (like Druckman and Lupia describe) or just attitude expression (in the form of survey responding)?
\item How important is information for opinion formation?
\end{itemize}
}

\againframe<4>{readings}

\frame{
\heading{Lau and Redlawsk}
\begin{itemize}\itemsep1em
\item How do people make decisions without full information?
\item Types of ``heuristics''
	\begin{itemize}
	\item Party
	\item Ideology
	\item Endorsements
	\item Viability/Horserace
	\item Appearance
	\end{itemize}
\end{itemize}
}


\frame{
\heading{Who will lead discussion next week?}
}

\appendix
\frame{}

\end{document}
