\documentclass[12pt]{beamer} %Makes presentation
\input{preamble}

\title{Preview of\\``How to measure opinions?''}

\date[]{September 25, 2013}

\begin{document}

\frame{\titlepage}

\frame{
\heading{Overview for next week}
}


\frame<1,4>[label=readings]{
\heading{Readings for Next Week}
\begin{itemize}\itemsep1em
\item<1> Don't read: Yeager et al. -- ``Measuring Americans' Issue Priorities''
\item<1-> Instead read: Tourangeau, Roger, and Kenneth A. Rasinski. 1988. ``Cognitive Processes Underlying Context Effects in Attitude Measurement.'' Psychological Bulletin 103 (3): 299--314.
\item<3-> Bishop et al. -- ``Opinions on Fictitious Issues''
\item<4-> Schuldt et al. -- `` `Global Warming' or `Climate Change' ''
\item<5-> Two optional readings
\end{itemize}
}

\againframe<2>{readings}

\frame{
\heading{Tourangeau and Rasinski}
\begin{itemize}\itemsep1em
\item How do they think people answer survey questions?
\item Is this consistent with our discussions from last week about the formation and structure of attitudes?
\item How does the context of a survey question influence how it is answered?
\item What impacts does their discussion have for interpreting the meaning of public opinion?
\end{itemize}
}

\againframe<3>{readings}

\frame{
\heading{Bishop et al.}
\begin{itemize}\itemsep1em
\item Why do people say they have opinions about issues?
\item How does the survey questionnaire and the survey interview shape when people report opinions?
\item Are there any ethical considerations related to this study?
\end{itemize}
}

\againframe<4>{readings}

\frame{
\heading{Schuldt et al.}
\begin{itemize}\itemsep1em
\item How much does question wording matter for what opinions people report on surveys?
\item When survey questions differ, are differences in opinions meaningful or artifacts of the survey context?
\item How can we know that what we're measuring are peoples' ``true'' opinions?
\end{itemize}
}

\againframe<5>{readings}

\frame{
\heading{Completely optional readings}
\begin{itemize}\itemsep1em
\item Yeager et al.: More about question wording
\item Krosnick et al.: ``No opinion'' response options
\end{itemize}
}


\frame{
\heading{Who will lead discussion next week?}
}

\appendix
\frame{}

\end{document}
