\documentclass[12pt]{beamer} %Makes presentation
\input{preamble}

\title{How do people form opinions?}

\date[]{September 18, 2013}

\begin{document}

\frame{\titlepage}


\frame{
\heading{Outline}
\begin{itemize}
\item Last week we talked about opinion in an abstract sense
\item This week we review that briefly
\item Then we'll talk about opinion in a concrete, psychological way
\end{itemize}
}


\section{Review}

\frame{
\heading{Review of Last Week}
\begin{itemize}\itemsep1em
\item ``Public opinion'' has many meanings
	\begin{itemize}
	\item Aggregation
	\item Majoritarian
	\item Discursive/Consensual
	\item Reification
	\end{itemize}
\end{itemize}
}

\frame{
\heading{Review of Last Week}
\begin{itemize}\itemsep1em
\item Democracy requires equal citizen participation
\item Citizens are the best judge of their own interests (?)
\item Competence is required for the right to participate
\end{itemize}
}

\frame{
\heading{Review of Last Week}
\begin{itemize}\itemsep1em
\item Rational government depends on:
	\begin{itemize}
	\item Pursuit of self-interest
	\item Well-specified preferences
	\end{itemize}
\end{itemize}
}

\frame{
\heading{Review of Last Week}
\begin{itemize}\itemsep1em
\item Four models of representation
	\begin{itemize}
	\item Promissory
	\item Anticipatory
	\item Gyroscopic
	\item Surrogate
	\end{itemize}
\end{itemize}
}


\section{New Material}

\frame{
\heading{Definitions}
\begin{itemize}
\item Attitude
\item Preference
\item Opinion?
\item<2> Do citizens have attitudes or preferences?
\end{itemize}
}

\frame{
\heading{Zaller's Model}
\begin{itemize}
\item Receive
\item Resist
\item Accept
\item Sample
\end{itemize}
}

\frame{
\heading{Processing Models}
\begin{itemize}
\item Memory-based (MB) Processing
\item Online (OL) Processing
\item<2-> Are MB and OL compatible?
\item<3-> Which seems more reasonable?
\end{itemize}
}

\frame{
\heading{Implications of Processing Models}
\begin{itemize}
\item Opinion stability
\item Susceptibility to influence
\item Answering survey questions
\end{itemize}
}

\frame{
\heading{Complications}
\begin{itemize}
\item What about affect?
\item<2-> Explicit and implicit attitudes
\item<3-> Do people have ``true'' attitudes?
\item<4-> Who appears knowledgeable?
\end{itemize}
}


\section{Attitudes}

\frame{
\heading{Attitudes}
\begin{itemize}
\item Structure
\item Function
\item Strength
\end{itemize}
}


\frame{
\heading{Attitude Structure}
\begin{itemize}
\item Intra-attitudinal structure
	\begin{itemize}
	\item Tripartite model
	\item Expectancy Value Model
	\item Belief Systems
	\item OL/MB differences?
	\end{itemize}
\item Inter-attitudinal structure
	\begin{itemize}
	\item More on this next week
	\end{itemize}
\end{itemize}
}

\frame{
\heading{Attitude Function}
\begin{itemize}
\item Object appraisal
\item Others?
	\begin{itemize}
	\item Knowledge/Frame of reference
	\item Instrumental/utilitarian
	\item Symbolic
	\item<2-> Value expression
	\item<2-> Social adjustment
	\item<2-> Ego defense
	\end{itemize}
\item<3> Most of these are just ``expressive''
\end{itemize}
}

\frame{
\heading{Attitude Strength}
\begin{itemize}
\item Types
	\begin{itemize}
	\item Experiential (meta-attitudinal)
	\item Operative
	\item Structural
	\end{itemize}
\item Impacts
	\begin{itemize}
	\item Stability
	\item Impact
	\end{itemize}
\end{itemize}
}

\frame{
\heading{Forms of Attitude Strength}
\begin{itemize}
\item Accessibility
\item Cognitive structure
\item Ambivalence
\item Importance
\item Certainty
\item Extremity?
\end{itemize}
}


\frame{
\heading{Information}
\begin{itemize}
\item Information
	\begin{itemize}
	\item<2-> Do citizens need information to form opinions?
	\item<2-> Does Dahl's idea of competence require information?
	\end{itemize}
\item Heuristics
	\begin{itemize}
	\item<3-> Do heuristics work in place of information?
	\end{itemize}
\end{itemize}
}

\section{}
\frame{}

\end{document}