\documentclass[12pt]{beamer} %Makes presentation
\input{preamble}

\title{Preview of\\``From Micro to Macro?''}

\date[]{November 6, 2013}

\begin{document}

\frame{\titlepage}

\frame{
\heading{Overview for next week}

Moving into the final section of the course...
}


\frame<4>[label=readings]{
\heading{Readings for Next Week}
\begin{itemize}\itemsep1em
\item Riker -- ``Liberalism Against Populism''
\item<2-> Page and Shapiro -- {\em The Rational Public}
\item<3-> Mackuen et al. -- ``Macropartisanship''
\item<4-> Druckman and Leeper -- ``Is Public Opinion Stable?''
\item<5-> Gilens -- ``Political Ignorance and Collective Policy Preferences''
\end{itemize}
}

\againframe<1>{readings}

\frame{
\heading{Riker}
\begin{itemize}\itemsep1em
\item Save this for next week!
\item You can read it now if you want, but we won't discuss it until next week.
\end{itemize}
}

\againframe<2>{readings}

\frame{
\heading{Page and Shapiro}
\begin{itemize}\itemsep1em
\item Just read Ch.10 (in the compendium)
\item Can we talk about public opinion in aggregate terms?
\item What empirical properties does ``macro'' opinion have?
\end{itemize}
}

\againframe<3>{readings}

\frame{
\heading{Mackuen et al.}
\begin{itemize}\itemsep1em
\item What is ``macropartisanship''?
\item How does it help us make sense of public opinion?
\item Can we generalize this concept to multi-party systems?
\end{itemize}
}


\againframe<4>{readings}

\frame{
\heading{Druckman and Leeper}
\begin{itemize}\itemsep1em
\item What properties does ``macro'' opinion have? What properties do ``micro'' opinions have?
\item Why do studies of macro and micro reach different conclusions about opinions?
\item What is opinion stability? And should we care about it?
\end{itemize}
}

\againframe<5>{readings}

\frame{
\heading{Gilens}
\begin{itemize}\itemsep1em
\item Following up on this week's discussion.
\item What is the impact of information on macro opinion?
\item Given what we know from this week about information and heuristics, should we trust these results?
\end{itemize}
}


\frame{
\heading{Who will write for next week?}
}

\appendix
\frame{}

\end{document}
